Component response rate variation drives stability in large complex systems

### 2018-05-19 ms.Rmd
More tweaks

To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from $2-32$ (see Supporting Information for different values of $\sigma$ and $C$). One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$ (note that under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$). I found that number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig. 3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity. For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supporting Information for full results). This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.

>>>>

To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from $2-32$ (see Supporting Information for different values of $\sigma$ and $C$). One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$ (note that under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$). I found that the number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig. 3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity. For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supporting Information for full results). This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.

Some care is needed when interpreting these results. First, I emphasise that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supporting Information). Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability. This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig. 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values. <!--- Mention specific ecological systems here? --->

<!--- Grob the genetic algorithm results and show that by changing the component response time, we can guide a system to much more stability. State that mutualism is not stabilised (SI), but this is unsurprising as mutualisms are unstable (cite Allesina). Finally, feasibility of ecological systems was not changed by $Var(\gamma)$. Conclude that more investigation is needed. Then need a strong concluding paragraph. --->
